---
title: "Running Multiple Mechanistic Niche Models"
author: "TP DuBose"
date: "7/14/2022"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse);library(sf);library(stars); library(maps); library(knitr)
```

First I identify study extent. I want to get a fine resolution that matches the microclimate data. So I use 

```{r}
se<-st_as_sf(map('state',c('mississippi', 'georgia',
             'florida','south carolina', 'north carolina', 'virginia',
             'alabama', 'tennessee'), fill=T, plot=F)) %>%
  st_make_valid() %>%
  st_transform(st_crs("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=km +no_defs"))
# create raster of study extent
se_r<-st_rasterize(se, dx=30, dy=30) %>% #turn into a raster
  st_as_sf(as_points = FALSE, merge = FALSE) %>%  #turn into a polygon
                       rowid_to_column() # add an id column
```


Next, I'm going to identify the points I want to run the models for. 
```{r}
test_spp<-c('Dryophytes chrysoscelis','Pseudacris crucifer','Lithobates catesbeianus', 'Anaxyrus fowleri', 'Gastrophryne carolinensis')
anuran_pts<-NULL
for(s in test_spp){
  anuran_pts<-bind_rows(anuran_pts,
                      read.csv(paste0('../National-RCS/data/occ_data/',s,'_20210330.csv')) %>% 
                                 dplyr::select(source, species, Longitude, Latitude,
                                               coordinatePrecision, elevation, year))
}
se_pts<-anuran_pts %>% st_as_sf(coords=c('Longitude','Latitude'), 
                                crs=st_crs("+proj=longlat +ellps=WGS84 +datum=WGS84")) %>%
  st_join(se_r %>% st_transform("+proj=longlat +ellps=WGS84 +datum=WGS84")) %>%
    filter(!is.na(rowid))  # remove points outside of our sample grid
pts_to_run<-NULL
for(i in sample(unique(se_pts$rowid),50)){
  pt_rw<-se_pts %>% filter(rowid==i)
  ptd<-pt_rw %>% st_distance() %>% rowSums()
  pts_to_run<-bind_rows(pts_to_run,
                        bind_cols(pt_rw[which(ptd==min(ptd)),] %>% filter(max(year)==year),
                                  t_dist=min(ptd),
                                  spp_all=paste(unique(pt_rw$species), collapse=', '),
                                  year_all=paste(unique(pt_rw$year), collapse=', ')))
}
pts_to_run %>% ggplot()+geom_sf(data=se_r %>% st_transform(st_crs(se_pts)), alpha=0.3) + 
  geom_sf(aes(color=species))+theme_classic()
write_sf(pts_to_run, 'microclimate_data/points_ran.shp', quiet=T)
se_pts_df1 <- pts_to_run %>% distinct(geometry, .keep_all=T) 
se_pts_df <- se_pts_df1 %>% st_coordinates() %>%
  bind_cols(rowid = se_pts_df1$rowid, spp_all=se_pts_df1$spp_all)
```

Next I bring in the traits we will use to create the models.

```{r}
traits<-read.csv('input_data/inputed_physio_traits.csv') %>%  #NEED TO ADD CODE HERE FOR TRAIT IMPUTAION
    left_join(read.csv('ATraiU 2.0/ATraiU2_summary_values_2022SEPT.csv') %>% 
              dplyr::select(species, nocturnal, diurnal, crepuscular) %>%
              mutate(across(-species, ~ifelse(.x > 0,1,0)))) %>%
  mutate(across(c(s.fossorial,s.arboreal), ~as.numeric(.x)))
kable(traits %>% sample_n(3), digits=2)
```

Then we need to chose a microclimate model as our input into the mechanistic niche model. To speed up the microclimate models, I use the code below, created by Mike Kearney and shared through the NicheMapR google group, to both download the NCEP data files and to convert it so that space is easier to query than time (source link)[https://groups.google.com/g/nichemapr/c/nOmFNhKTFrM/m/xTNL-GJ_AQAJ]. 

```{r download ncep, eval=F}
# Code by Mike Kearney!
library(curl)
years <- seq(2019,2021)
vars <- c('tmax.2m.gauss.','dswrf.sfc.gauss.', 'prate.sfc.gauss.', 'vwnd.10m.gauss.',
          'uwnd.10m.gauss.', 'air.2m.gauss.', 'tmin.2m.gauss.', 'shum.2m.gauss.',
          'pres.sfc.gauss.', 'dlwrf.sfc.gauss.', 'ulwrf.sfc.gauss.', 'tcdc.eatm.gauss.')
for(year in years){
  for(j in 1:length(vars)){
    curl_download(paste0("ftp://ftp.cdc.noaa.gov/Datasets/ncep.reanalysis2/gaussian_grid/",vars[j],year, ".nc"), 
                  paste0("../ncep_data/",vars[j],year, ".nc"),quiet=TRUE, mode="wb")   
  }
}

library(ncdf4)
dir2do <- "../ncep_data/"
setwd(dir2do)
files<-list.files()
files<-files[grep(".nc", x = files)]
vars4 <- c("tmin", "tmax", "air", "shum", "uwnd", "vwnd")
vars3 <- c("pres", "tcdc", "dswrf", "dlwrf", "ulwrf", "prate")

allvars <- c(vars4, vars3)

for(i in 1:length(files)){
  filenm <- files[i]
  orig <- nc_open(filenm)
  t<-ncvar_get(nc = orig, varid = "time", start = 1, count = -1)
  
  varid <- names(orig$var)[which(names(orig$var) %in% allvars)]
  
  if(varid %in% vars3){
    start <- c(1,1,1) # compose the start location vector for getting the data out of the ncdf file for this location
    count <- c(-1,-1,-1) # how much data are wanted? (-1 means all)
  }else{
    start <- c(1,1,1,1) # compose the start location vector for getting the data out of the ncdf file for this location
    count <- c(-1,-1,-1,-1) # how much data are wanted? (-1 means all)
  }
  
  data <- as.numeric(ncvar_get(orig, varid = varid, start = start, count)) # get the data
  lon<-ncvar_get(nc = orig, varid = "lon", start = 1, count = -1) # longitudes
  lat<-ncvar_get(nc = orig, varid = "lat", start = 1, count = -1) # latitudes
  t2 <- ncdim_def(name = "time", units = " hours since 1800-1-1 00:00:0.0", vals = t, unlim=FALSE) # define a time variable
  lon2 <- ncdim_def(name = "lon", units = "degrees_east", vals = lon, unlim=FALSE) # define a longitude variable
  lat2 <- ncdim_def(name = "lat", units = "degrees_north", vals = lat, unlim=FALSE) # define a latitude variable
  att<-ncatt_get(orig, varid, attname=NA, verbose=FALSE) # get attributes from original file
  if(varid %in% vars4){
    level<-ncvar_get(nc = orig, varid = "level", start = 1, count = -1) # levels
    level2 <- ncdim_def(name = "level", units = "m", vals = level, unlim=FALSE) # define a latitude variable
  }
  
  
  if(varid %in% vars3){
    var1 <- ncvar_def(name = varid, units = att$units, dim = list(lon2,lat2,t2), missval = att$missing_value, longname = att$long_name, prec='float') # define the variable
  }else{
    var1 <- ncvar_def(name = varid, units = att$units, dim = list(lon2,lat2,level2,t2), missval = att$missing_value, longname = att$long_name, prec='float') # define the variable
  }
  
  vars <- list(var1) # get variable definition
  file2<-gsub('.nc', replacement = '_time2.nc', x = filenm) # make a new file name
  file3<-gsub('.nc', replacement = '_time.nc', x = filenm) # make a new file name
  ncnew <- nc_create(file2,vars) # create a new ncdf file to put the data in
  ncvar_put(ncnew,var1,data) # put the data in
  nc_close(ncnew) # close the new file
  if(varid %in% vars3){
    cmd<-paste0("ncpdq -a lon,lat,time ",file2," ",file3) # command for ncpdq - permute to reverse order
  }else{
    cmd<-paste0("ncpdq -a lon,lat,level,time ",file2," ",file3) # command for ncpdq - permute to reverse order
  }
  
  system(cmd) # run ncpdq
  
  file.remove(file2) # delete the temporary file
}
```

We then need to run microclimate and ectotherm models for many points across the landscape. Ideally this will be at the resolution of the environmental datasets. The NicheMapR team has explored how ordinary kringing can be used to interpolate microclimate data to a physiological parameter (article link here)[https://onlinelibrary-wiley-com.ezproxy.lib.vt.edu/doi/10.1111/jbi.13152]. In 2020, there was a discussion of using rasters to speed processing for large geographic extents (thread here)[https://groups.google.com/g/nichemapr/c/FNzXsigRdmk/m/7z_4-rBaBgAJ]

So I will want to run a lot of ectotherm models to translate microclimate data and species specific traits into estimated body temperature. To make this more legible, I create the function below to wrap around ectotherm and output the data we are interested in.

```{r}
Te_est<-function(spp, save=F){
  spp_traits <- traits %>% filter(species == spp)
    ecto<-ectotherm(Ww_g=spp_traits["Mass"], 
                    CT_max=spp_traits["CTmax"], 
                    CT_min=spp_traits["CTmin"], 
                    T_pref=spp_traits["Tpref"],
                    diurn=spp_traits["diurnal"],
                    nocturn=spp_traits["nocturnal"],
                    crepus=spp_traits["crepuscular"],
                    burrow=spp_traits["s.fossorial"],
                    climb=spp_traits["s.arboreal"],
                    shdburrow = 1, #the animal's retreat is in the open (0), in the shade when above or below CTmin in sun (1) or in shade always
                    maxdepth = 10, #maximum depth of the burrow
                    T_F_min=max(c(19, as.numeric(spp_traits["Tmerge"]))),
                    T_F_max=min(c(35, as.numeric(spp_traits["CTmax"]))),
                    T_RB_min=spp_traits["Tmerge"],
                    T_B_min=spp_traits["Tmerge"])
    if(save==T){saveRDS(ecto,
                paste0('microclimate_data/pt',rw,'/!',spp,'.rds'))}
    body_temps<-ecto$environ[,1:5]
    output<-body_temps %>% as_tibble() %>%
      mutate(hot_tsm=as.numeric(spp_traits["CTmax"])-TC) %>% 
      group_by(DOY) %>% 
      mutate(dayrange=max(TC)-min(TC)) %>%
      ungroup() %>%
      summarize(n_total=n(),
                n_above=sum(hot_tsm < 0),
                max_TC=max(TC),
                tsm=min(hot_tsm),
                max_dayrange=max(dayrange),
                avg_dayrange=mean(dayrange),
                X=micro$longlat[1],
                Y=micro$longlat[2],
                species=spp)
return(output)
}
Te_est("Pseudacris crucifer")
```

So then we can start with the simplest comparison: at one location where a physiological trait was determined. 

# One Point Interspecific Comparison

# Intraspecific Comparisons

```{r}
ecto_out<-NULL
for(rw in 41:nrow(se_pts_df)){
  #dir.create(paste0('microclimate_data/pt',rw))
  micro<-micro_ncep(loc = c(as.numeric(se_pts_df[rw,1]),as.numeric(se_pts_df[rw,2])),
                    dstart = "01/01/2020", dfinish = "31/12/2020",
                    dem.res=1000, # requested resolution of the DEM from elevatr in m
                    DEP = c(0, 2.5,  5,  10,  15,  20,  30,  50,  100,  200), #cm
                    minshade = 0, maxshade = 90, runshade=1,
                    write_input=0, #save=1,
                    soilgrids=1, #CHECK WHY THIS ISN'T WORKING
                    spatial='../ncep_data',
                    Usrhyt = 0.01) # local height for organism
  #file.copy(list.files(pattern='.Rda'), 
  #          paste0('microclimate_data/pt', rw, '/',list.files(pattern='.Rda')))
  
  for(spp in test_spp){
    if(grepl(spp, se_pts_df[rw,4])){
      ecto_out<-bind_rows(ecto_out,
                          Te_est(spp) %>% mutate(rowid=rw))
    print(paste('done with ',spp,' point', rw, 'of', nrow(se_pts_df)))
    }
  }
}
```

```{r ectotherm function for loop, eval=T}
# ectotherm for loop shown here.
```


```{r}
ecto_out<-NULL
for(rw in 1:10){
for(k in test_spp){
  if(k %in% gsub('!','', gsub('.rds','', 
                              list.files(paste0('microclimate_data/pt',rw,'/'), pattern='.rds')))){
  e1<-readRDS(paste0('microclimate_data/pt', rw, '/!', k, '.rds'))
  body_temps<-e1$environ[,1:5]
  ecto_out<-bind_rows(ecto_out,
                        body_temps %>% as_tibble() %>%
                          mutate(hot_tsm=as.numeric(spp_traits["CTmax"])-TC) %>% 
                          group_by(DOY) %>% 
                          mutate(dayrange=max(TC)-min(TC)) %>%
                          ungroup() %>%
                          summarize(n_total=n(),
                                    n_above=sum(hot_tsm < 0),
                                    max_TC=max(TC),
                                    tsm=min(hot_tsm),
                                    max_dayrange=max(dayrange),
                                    avg_dayrange=mean(dayrange),
                                    X=as.numeric(va_pts_df[rw,1]),
                                    Y=as.numeric(va_pts_df[rw,2]),
                                    species=k,
                                    rowid=rw))
  }
}
}
ecto_out
length(unique(ecto_out$rowid))
```

```{r}
ecto_out %>%
  group_by(species) %>%
  summarize(min(tsm),
            max(max_TC),
            sum(n_above),
            mean(n_above)) %>%
  arrange(`min(tsm)`)

```


```{r}
ecto_out %>% distinct(X, Y, species, .keep_all=T) %>%
  st_as_sf(coords=c('X','Y'),
           crs=st_crs("+proj=longlat +ellps=WGS84 +datum=WGS84")) %>%
  ggplot()+geom_sf(data=se_r, fill=NA)+
  geom_sf(aes(fill=tsm), pch=21, size=2.5)+
  facet_wrap(~species)+theme_classic()+
  scale_fill_gradient2('Warming\nTolerance')
ggsave('results/thermal safety margin space.jpg', width=6.5, height=3)

library(cowplot)
plot_grid(ecto_out %>%
  group_by(rowid, X, Y) %>%
  summarize(n_spp=n_distinct(species),
            wt_mean=mean(tsm),
            wt_sd=sd(tsm)) %>%
      st_as_sf(coords=c('X','Y'),
           crs=st_crs("+proj=longlat +ellps=WGS84 +datum=WGS84")) %>%
  ggplot()+
  geom_sf(aes(color=n_spp, geometry=geometry))+
  scale_color_viridis_c('n\nspecies')+
    theme_void(),
  ecto_out %>%
 group_by(rowid, X, Y) %>%
  summarize(n_spp=n_distinct(species),
            wt_mean=mean(tsm),
            wt_sd=sd(tsm)) %>%
      st_as_sf(coords=c('X','Y'),
           crs=st_crs("+proj=longlat +ellps=WGS84 +datum=WGS84")) %>% 
  ggplot()+
  geom_sf(aes(color=wt_mean, geometry=geometry))+
  scale_color_gradient2('Warming\nTolerance')+
    theme_void())

ggsave('results/assemblage tolerance space.jpg', width=6, height=3)
```

```{r}
ecto_out %>% group_by(species) %>%
  mutate(ptdist=max(`Y`)-`Y`) %>%
  ggplot()+
  geom_hline(yintercept=0)+
  geom_smooth(aes(x=Y, y=tsm, color=species, fill=species), method='lm', level=.8)+
  geom_point(aes(x=Y, y=tsm, color=species), size=2, alpha=.5)+
  labs(y='Warming Tolerance (deg C)',x='Latitude (deg))')+
  theme_classic()+
  theme(axis.title = element_text(size=8),
        axis.text=element_text(size=8))
ggsave('results/tsm_lat_gradient2_real.jpg', width = 5, height=3)

```

# Do estimated microclimates differ within the climate data resolution?

Below I run two microclimate models using a DEM resolution of 1km and a climate layer resolution of 2.5 degrees at 6hr increments. Will they differ because of the DEM resolution? Answer is yes!

```{r}
se_pts<-se_r %>% 
  st_transform("+proj=longlat +ellps=WGS84 +datum=WGS84") %>% 
  st_centroid() %>% st_coordinates()
m1<-micro_ncep(loc = c(as.numeric(se_pts[1,1]),
                       as.numeric(se_pts[1,2])),
                    dstart = "01/01/2020", dfinish = "31/12/2020",
                    dem.res=1000, # requested resolution of the DEM from elevatr in m
                    DEP = c(0, 2.5,  5,  10,  15,
                            20,  30,  50,  100,  200), #cm
                    minshade = 0, maxshade = 90, runshade=1,
                    write_input=0, save=0,
                    soilgrids=1,
                    spatial='../ncep_data',
                    Usrhyt = 0.01)
m2<-micro_ncep(loc = c(as.numeric(se_pts[2,1]),
                       as.numeric(se_pts[2,2])),
                    dstart = "01/01/2020", dfinish = "31/12/2020",
                    dem.res=1000, # requested resolution of the DEM from elevatr in m
                    DEP = c(0, 2.5,  5,  10,  15,
                            20,  30,  50,  100,  200), #cm
                    minshade = 0, maxshade = 90, runshade=1,
                    write_input=0, save=0,
                    soilgrids=1,
                    spatial='../ncep_data',
                    Usrhyt = 0.01)
identical(m1, m2)
```

```{r temp plots within climate grid, echo=F}
bind_rows(m1$metout %>% as_tibble() %>% mutate(name='point 1') %>%rowid_to_column(),
          m2$metout %>% as_tibble() %>% mutate(name='point 2') %>%rowid_to_column()) %>%
  filter(DOY==1) %>%
  ggplot()+
  geom_line(aes(x=TIME, y=TAREF, color=name), alpha=0.5)+
  scale_color_viridis_d('Points')+
  scale_y_continuous('Temperature at 1m')+
  scale_x_continuous('Time (minutes)')+
  theme_classic()+
  theme(legend.position = c(.9, .9))+
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

library(cowplot)
plot_grid(st_sfc(st_point(c(as.numeric(se_pts[1,1]), as.numeric(se_pts[1,2]))),
          st_point(c(as.numeric(se_pts[2,1]), as.numeric(se_pts[2,2]))),
       crs=4326) %>%
  st_transform(st_crs("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=km +no_defs")) %>%
  ggplot()+geom_sf(data=se_r[1:5,])+geom_sf()+theme_classic(),
  ggplot()+geom_sf(data=se[5,])+geom_sf(data=se_r[1:5,], fill='red',color='red')+theme_classic(),
  nrow=1)
```

# Decisions that remain:

1) which climate model to use
2) temporal resolution to use [ year that matches observation or single year across all observations]
3) spatial resolution to use 

today (8/31) I'm going to attend arc office hours to see if I can NicheMapR onto the super computer and then maybe get advice for parallelizing the model creation step.

