---
title: "5_intraspecific variation"
author: "Traci DuBose"
date: "1/24/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
tmplab<-function(x){paste0(x, '°C')}
latlab<-function(x){paste0(x, '° N')}
```

# Investigate how traits affect our understanding of warming tolerance

Using either the least sensitive or most sensitive traits, I build "trait monsters", or species with artificial combinations of traits, to investigate the boundaries of warming tolerance built using MNMs. These trait monsters represent fake animals that have artificial combinations of traits but might represent our estimates of the warming tolerance of the least sensitive animal and the warming tolerance of the most sensitive animal. 

```{r load data}
in.traits<-read.csv('input_data/intraspecific_variation_subset_APPENDIX.csv')
in.ecto<-read.csv('intermediate_results/intra_summarized_bodytemps_230127.csv')
in.pts<-read.csv('intermediate_results/intra_microclimate_info_230127.csv')
pts<-read.csv('intermediate_results/points_ran_2023-01-10.csv') 
pts.res<-read.csv('results/pt_wtmin_sum.csv')
#sample size possible
pts_poss<-pts %>% select(X, Y, rowid, mnm_spp)%>%
  separate(mnm_spp, into=paste0('sp', 1:13), sep=', ') %>%
  pivot_longer(starts_with('sp')) %>%
  filter(!is.na(value)) %>%
  filter(value %in% unique(in.ecto$species)) %>%
  distinct(rowid, value) %>%
  count(value)
```

I ran these trait monsters for a subset of all possible points. Below describes how well the entire occupied grid was sampled. 

```{r}
in.ecto %>%
  distinct(species, rowid) %>%
  left_join(in.pts, by='rowid') %>%
  distinct(species, rowid, lat) %>%
  group_by(species) %>%
  summarize(n=n(),
            minlat=min(lat), maxlat=max(lat)) %>%
  left_join(pts_poss %>% rename(species=value, n_possible=n), 
            by='species') %>%
  mutate(per.pts=round(n/n_possible*100, 2))
```


```{r}
ss.comp<-in.ecto %>%
  filter(src %in% c('super','sensitive')) %>%
  left_join(in.pts, by='rowid') %>%
  distinct(species, rowid, lat, src, .keep_all = T) %>%
  mutate(nsp=paste(src, species))
```

We can recreate Figure 1 in the manuscript but also plot the sensitive and super trait monsters, when available. 

```{r}
# rank wt
wtord <- pts.res %>% group_by(species) %>% 
  arrange(medWT) %>% pull(species)

ss.comp %>%
   pivot_longer(cols=c('WT_min25','WT_ptmin', 'WT_min50')) %>%
  mutate(nm.F=factor(name, levels=c('WT_ptmin','WT_min25','WT_min50'),
                     labels=c('minimum','25th quartile','median'))) %>%
   group_by(species, src, nm.F) %>% 
  summarize(medWT=median(value),
            WTmin2=quantile(value, .025),
            WTmin97=quantile(value, .975), .groups='drop') %>%
  bind_rows(pts.res %>% mutate(src='interspecific')) %>%
  mutate(nm.F=factor(nm.F, levels=c('minimum','25th quartile','median'))) %>%
  mutate(spF=factor(species, levels=rev(wtord))) %>%
  ggplot()+
  geom_hline(yintercept=0, linetype='dashed', color='grey')+
  geom_linerange(aes(x=spF, ymin=WTmin2, ymax=WTmin97, color=src),
             position=position_dodge(width=.5))+
  geom_point(aes(x=spF, y=medWT, color=src), size=1.7,
             position=position_dodge(width=.5))+
  scale_y_continuous('Minimum Warming Tolerance', labels = tmplab)+
  scale_color_manual('Trait Type', values=c('black','blue','red'))+
  theme_classic()+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_text(face='italic'),
        legend.text = element_text(size=8),
        legend.title = element_text(size=8),
        legend.position='bottom')+
  coord_flip()+
  facet_grid(~nm.F)
ggsave('results/FigS3_1_rankwarmingtolerance.jpg', width=6.5, height=4)

```

Below is code to compare the relationship between warming tolerance and latitude between trait monsters of the same species. 

```{r}
library(lme4)
wtmin.mod<-lmer(WT_ptmin~lat:species + species + 0 + (1|src), data=ss.comp)
(mod.sum<-summary(wtmin.mod))
wtmin.pred<-cbind(ss.comp, fit=predict(wtmin.mod))
ggplot(wtmin.pred)+
  geom_point(aes(x=lat, y=WT_ptmin, color=src), alpha=0.02)+
  geom_line(aes(x=lat, y=fit, color=src))+
  facet_grid(~species)+
  theme_classic()+
  theme(legend.position='bottom')
```

And then we can look to see how the relationship between warming tolerance and latitude changes between the sensitive and super trait monsters. 

```{r}
ss.comp %>%
  pivot_longer(cols=c('WT_min25','WT_ptmin', 'WT_min50')) %>%
  mutate(nm.F=factor(name, levels=c('WT_ptmin','WT_min25','WT_min50'),
                     labels=c('minimum','25th quartile','median'))) %>%
  ggplot()+
  geom_hline(yintercept=0, color='grey')+
  geom_smooth(aes(x=lat, y=value, color=src), method='lm',formulat='y~x')+
  facet_grid(nm.F~species)+
  theme_classic()
```
