---
title: "NicheMapper Sensitivity Analysis"
author: "TP DuBose & Vj Catalan"
date: "6/29/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages('NicheMapR')
library(NicheMapR); library(tidyverse)
library(knitr)
library(lubridate) #to deal with day times when plotting
```

## How sensitive is the ectotherm model to changes in emergence temperature

What is our predicted value we care about? Body temperature shifts through the day (expected). We will look at two response variables in the model that we've discussed as important for anuran's response to climate change: activity (ACT column in environ output) and body temperature. There are many ways we could quantify body temperature. We could look at proximity to CTmax, time above CTmax, or mode body temperature (does it get to regulate to its Tpref often?). I'll start with activity since it is easier to think about. 

```{r}
longlat <- c(-80.41, 37.23)
micro <- micro_global(loc = longlat,
                      timeinterval = 12)
```

Below I wrote a for loop to start~

```{r ofat analysis}
all_results<-NULL
k=13
for(k in 13:19){
  # run the model with the new tmerge
  ecto<-ectotherm(Ww_g=6.6, 
                  CT_max=32.2, 
                  CT_min=0, 
                  T_pref=23.9,
                  diurn=0,
                  nocturn=1,
                  crepus=1,
                  burrow=1,
                  climb=0,
                  shdburrow = 1, #the animal's retreat is in the open (0), in the shade when above or below CTmin in sun (1) or in shade always
                  maxdepth = 3, #maximum depth of the burrow
                  T_B_min=k,
                  T_RB_min=k)
  # pull results out of the model
  new_results<-ecto$environ %>% as_tibble() %>%
    mutate(Tmerge=k)
  # save the changes in the outcome
  all_results<-bind_rows(all_results, new_results)
}
```

```{r results table, echo=F}
kable(all_results %>% count(Tmerge,ACT) %>% 
        pivot_wider(names_from=ACT, values_from=n) %>%
        rename(inactive=`0`, basking=`1`))
```
So summarizing our results

```{r}
all_results %>%
  count(Tmerge, ACT) %>%
  mutate(behavior=recode(ACT, `0`='inactive',`1`='active')) %>%
  ggplot()+
  geom_point(aes(x=Tmerge, y=n))+
  facet_wrap(~behavior, scales='free_y')+
  theme_bw()+
  labs(main='Point-wise Sensitivity to Emergence Temp',
       x='Emergence Temperature', y='Hours across a year')
ggsave('sensivity_ex_plot.jpg', width=3, height=2)
```

```{r}
View(body_temp)
body_temp<-as_tibble(all_results) %>%
  mutate(date_x=as_date(DOY-1, origin='2019-01-01'),
         day_time=ymd_h(paste0(date_x, " ", TIME))) 
#body_temp %>% select(YEAR, DOY, TIME, date_x, day_time)
#body_temp %>% pull(date_x) %>% unique()
body_temp %>%
  filter(day_time<'2019-09-01 0:00' & day_time > '2019-06-01 0:00') %>%
ggplot()+
  geom_point(aes(x=day_time,y=TC, shape=as.character(ACT)))+
  geom_line(aes(x=day_time,y=TC, 
                color=as.factor(Tmerge), group=as.factor(Tmerge)),
            size=1.5, alpha=0.2)+
  #facet_wrap(~DOY, nrow=1)+
  ggtitle(paste(longlat[1],longlat[2], sep=', '))+
  scale_color_viridis_d(name='Tmerge')+
  scale_y_continuous('Body Temperature')+
  scale_shape('Activity')+
  theme_bw()+
  theme(axis.text.x=element_text(angle=30, size=6, hjust=.9))+
  facet_wrap(~date_x, scales='free_x', nrow=1)
```

# Global Sensitivity Analysis

TPD is currently reading Scalli 2008, an overview of global sensitivity analysis. Below, the code just follows our inspiration article. From the sensitivity package: soboljansen implements the Monte Carlo estimation of the Sobol’ indices for both first-order and
total indices at the same time (alltogether 2p indices), at a total cost of (p+2)×n model evaluations.

The model input into the function needs to either have a predict function or produce a matrix or 3d array as its output (none of which the ectotherm model does). So I'll need to make a function that converts the ectotherm model into the output we want to analyze. 

Satalli et al 2010 notes - 
individual factos (S_i) and total factor (S_Ti) can be defined in terms of expected variance reduction. Vx(Ex~i(Y|Xi)) is the variance reduction if Xi could be fixed while Ex~1(VXi(Y|X~i)) is the expected variance if everything BUT Xi could be fixed. So they represent the most and least marginal variances respectively -- bounds on our variance from the model!

To compute this, you use two matrices (nboots x factors changing in model). During the computations, one matrix has all factors varying except one (which is pulled from the second matrix and considered fixed). 

Si = normalized index of the variance based first order effect

STi = total effect of factor Xi

TO DO:
  look at Scalli 2008 to get units of S
  get soboljansen to run (currently getting datafram errors, takes a long time to run)
  look at rules for random matrixes to calculate Jansen indices
  
So to start, we set up our random matrices. They need to be number of samples x number of parameters large. Current matrix set up to sample from a minimum value, maximum value, and resolution the number of samples we want to test for two parameters. 

```{r}
#install.packages('sensitivity')
library(sensitivity)
n.samp=2
for(rdf in c('X1', 'X2')){
  
assign(rdf,
       data.frame(mass=sample(seq(.5,10,by=0.1),size=n.samp),
               ctmax=sample(seq(30,39,by=0.5), size = n.samp),
               ctmin=sample(seq(-5,5,by=0.5), size = n.samp)))
}
```

Below is the function we use to assess sensitivity. We set the parameters for the ectotherm model. And then we define which response variable we are interested in.

```{r}
ecto.sen<-function(x){
  #run the model
  ecto<-ectotherm(Ww_g=x[,1], 
            CT_max=x[,2], 
            CT_min=x[,3], 
            T_pref=23.9,
            diurn=0,
            nocturn=1,
            crepus=1,
            burrow=1,
            climb=0,
            shdburrow = 1)
  #the response variable we are interested in
  main_results<-ecto$environ
  res<-length(main_results[,9]) - sum(main_results[,9]) 
  as.numeric(res)
}

# sensitivity analysis
x <- soboljansen(model = ecto.sen, 
                 X1, #first random sample
                 X2 #second random sample
)
#3 rows in random matrix,3 nboot = 2059 seconds
#df error seems to go away when n.samp = 10
# its in the tell function, which I can't seem to find
#10 rows no bootstrap = 5055 seconds
# three variables, three samples : 148586 seconds

```

The outputs will be the first order (variance due to that parameter changing) and total (variance due to that parameter (plus other parameters changing)) indices. 

```{r}
# first order
x$S
# third order
x$T
```



