---
title: "Identifying Focal Taxa"
author: "TP DuBose"
date: "September 2020"
output: pdf_document
---

The goal of this script was to identify focal taxa to search for that spanned a gradient of rare to common species. 

```{r setup, include=FALSE}
library(tidyverse); library(sf); library(knitr)
```

First, I brought in a preliminary calculation of RCS index values. Then I brought in the point occurrences I used to generate those RCS values and created a spatial dataframe object of those occurrences.

```{r}
rcs_index<-read.csv('C:/Users/Owner/Desktop/Git/National-RCS/rcs_results/RCS_table_20210405.csv') %>%
  dplyr::select(species, RCS_WS, RCS_buff)

all_anuran_occ<- read.csv('C:/Users/Owner/Desktop/Git/National-RCS/data/old_files/anuran_occ_all20200708.csv')
nrow(all_anuran_occ)
head(all_anuran_occ)
#count of all obs
anuran_count<-all_anuran_occ %>% count(final.taxa) %>%
  rename(total.obs='n')

# spatial dataframe of occurrences
crs.geo <- st_crs("+proj=longlat +ellps=WGS84 +datum=WGS84")
all_anuran_sf<-st_as_sf(x=all_anuran_occ, 
                        coords=c("Longitude","Latitude"), 
                        crs=crs.geo)
```

# modified SE USGS region

I then created a polygon that had the states of united states from the *maps* package. I restricted the states I considered the Southeast United States to Mississippi, Georgia, Florida, South Carolina, North Carolina, Virginia, Alabama, and Tennessee. This is slightly modified from the USGS Southeast (or South and East) region: lost IA, MO, TX & OK but gained WV & VA.

```{r}
library(ggplot2); library(maps)
states<-st_as_sf(map('state', fill=T))
st_crs(states)==crs.geo

SE_states<-c('mississippi', 'georgia',
             'florida','south carolina', 'north carolina', 'virginia',
              'alabama', 'tennessee')
```

I then spatially joined the occurrence dataset with the states to identify where each occurrence landed on a map. Then, I filtered the dataset to only those within Virginia and classified anuran rarity based on its RCS value. I did this to try to get a sense of the availability of data among different species vs. their rarity/sensitivity. 

```{r}
sf::sf_use_s2(FALSE)
state_anurans<-st_join(all_anuran_sf, 
                       states %>% st_buffer(0) %>% st_make_valid())
sf::sf_use_s2(TRUE)
(sp_choice_list<-state_anurans %>% 
  rename(state='ID.y') %>% filter(state == "virginia") %>%
  count(final.taxa) %>% 
  as_tibble() %>% dplyr::select(-geometry) %>%
  left_join(anuran_count, by='final.taxa') %>%
  mutate(percent.obs=(n/total.obs)*100) %>%
  arrange(desc(percent.obs)) %>%
  left_join(rcs_index, by=c('final.taxa'='species')) %>%
    mutate(select_group=case_when(RCS_WS <0.575 ~ 'not vulnerable',
                                  RCS_WS >=0.575 & RCS_WS < 0.75 ~ 'moderate',
                                  RCS_WS >=0.75 & RCS_WS < 0.825 ~ 'vulnerable',
                                  RCS_WS >=0.825 & RCS_WS <=1 ~ 'very vulnerable')))
sp_choice_list %>% filter(select_group=='not vulnerable') 
```
```{r, echo=F}
ggplot()+geom_histogram(data=sp_choice_list, aes(x=RCS_WS, fill=select_group),
                        binwidth=0.05)
```

Using that species list, I sampled 4 species from each of my sensitivity classifications. These 4 species would be a starting list for our focal taxa. 

```{r}
sp_choice_list %>% count(select_group)
set.seed(10403)
(sampled_spp<-sp_choice_list %>% 
  group_by(select_group) %>%
  sample_n(size=4, replace=F) %>%
  arrange(desc(RCS_WS)))
```

Below is a list of the first 16 taxa we used as our focal list. This list grew as Vj completed taxa and incorporates the taxa searched before 2020. 

```{r, echo=F}
mnm_subset<-c('Lithobates capito', 'Pseudacris brimleyi', 
              'Pseudacris ornata', 'Lithobates grylio',
              'Pseudacris fouquettei', 'Pseudacris ocularis',
              'Pseudacris brachyphona', 'Anaxyrus quercicus',
              'Lithobates areolatus', 'Scaphiopus holbrookii',
              'Anaxyrus punctatus', 'Pseudacris triseriata',
              'Pseudacris nigrita', 'Lithobates catesbeianus',
              'Dryophytes versicolor', 'Anaxyrus americanus')
kable(sp_choice_list %>% filter(final.taxa %in% mnm_subset) %>%
  arrange(RCS_WS))
```

Below are some plots I used to investigate different anuran distributions. 

```{r, echo=F}
ggplot()+
  geom_sf(data=states)+
  geom_sf(data=state_anurans[state_anurans$final.taxa=='Dryophytes versicolor',])+
  ggtitle('Dryophytes versicolor')
ggplot()+
  geom_sf(data=states)+
  geom_sf(data=state_anurans[state_anurans$final.taxa=='Anaxyrus americanus',])+
  ggtitle('Anaxyrus americanus')
ggplot()+
  geom_sf(data=states)+
  geom_sf(data=all_anuran_sf[all_anuran_sf$final.taxa=='Anaxyrus californicus',])+
  ggtitle('Anaxyrus californicus')
```

I used the below code to investigate the distribution of other southeastern anuran taxa. It outputs a pdf that contains all the occurrences for taxa found within the states set as our region above.  

```{r}
sp_choice_list$final.taxa
sp_lists_jan<-state_anurans %>% 
  rename(state='ID.y') %>% filter(state %in% SE_states) %>%
  group_by(final.taxa) %>% slice(1) %>%
  select(family, genus, final.taxa, state)
pdf('se_anuran_list.pdf')
for(i in sp_lists_jan$final.taxa){
  print(state_anurans %>% filter(final.taxa==i) %>%
    ggplot()+geom_sf(data=states)+geom_sf()+ggtitle(i))
}
dev.off()
```